---
title: "Data Science Salary Analysis"
author: "Sue Susman, MEd, BSN, RN"
date: "`r Sys.Date()`"
output: 
  pdf_document: 
    toc: yes
    toc_depth: 6
    fig_caption: yes
    number_sections: yes
    keep_tex: yes
    fig_width: 10
    fig_height: 8
  word_document: 
    toc: yes
    fig_caption: yes
    keep_md: yes
    toc_depth: 6
    fig_width: 10
    fig_height: 8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Project Overview:

The CEO of the small, but rapidly expanding company I work for is interested in adding a full-time data scientist to the staff. While the company is in the US, there is a possibility for this position to be remote for the right person. Factors to consider include rising salaries in the recession, along with a highly competitive job market. My research will explore salaries of data jobs in and out of the US. I will attempt to identify trends and patterns in the data science job market, including experience levels, location of companies, company sizes, salary differences between hiring someone in the US vs. someone outside of the US.

```{r}
options(repos = "https://cran.rstudio.com/")
```

```{r Importing and Loading Packages}
    install.packages("tidyverse")
    install.packages("ggwordcloud")
    install.packages("viridis")
    install.packages("scales")
    install.packages("dplyr")

    library(tidyverse)
    library(ggwordcloud)
    library(viridis)
    library(scales)
    library(dplyr)

    options(repr.plot.width = 15, repr.plot.height = 15)
```

```{r Loading the Datasets}
#Main Data set
salary_df <- read.csv("data.csv")

#Country codes Data set, used for cleaning later
iso_df <- read.csv("wikipedia-iso-country-codes.csv")
```

Checking the dimensions of a dataframe can be useful for understanding the size and structure of your data and for performing various operations and analyses based on the number of rows and columns present.

```{r Inspecting the Data}
t(t(names(salary_df))) # In summary, the code transposes the column names of the data_df dataframe and returns them as a column matrix to display the column names in a user friendly manner.
```
```{r}
#Dimensions of the data frame
data_dimensions <- dim(salary_df)
num_rows <- data_dimensions[1]
num_cols <- data_dimensions[2]

text <- paste("There are", num_rows,"Rows", "and", num_cols, "Columns", "in this dataset")

print(text)
```

```{r}
head(salary_df)  #Displays the first couple of rows of the data frame
```

```{r}
str(salary_df)  #Display list of columns and data types
```


```{r}
summary(salary_df)  #Statistical summary of data 
```

```{r}
# Checking for any NA values
any(is.na(salary_df))
```
```{r}
#Checking for unique values per column
library(dplyr)
salary_df %>% 
  summarise(
    work_year = n_distinct(work_year),
    experience_level = n_distinct(experience_level),
    employment_type = n_distinct(employment_type),
    job_title = n_distinct(job_title),
    salary = n_distinct(salary),
    salary_currency = n_distinct(salary_currency),
    salary_in_usd = n_distinct(salary_in_usd),
    employee_residence = n_distinct(employee_residence),
    remote_ratio = n_distinct(remote_ratio),
    company_location = n_distinct(company_location),
    company_size = n_distinct(company_size)
  )
```

Cleaning the Data
There are a few problems we need to fix:
1. The columns experience_level, employment_type, employee_residence, company_size and company_location contain abbreviations that may not be understood to some, we will rename the data contained in those columns to more descriptive labels.¶

```{r}
#Renamed abbreviations for the experience_level column
salary_df$experience_level[salary_df$experience_level == "SE"] <- "Senior"
salary_df$experience_level[salary_df$experience_level == "MI"] <- "Indermediate"
salary_df$experience_level[salary_df$experience_level == "EN"] <- "Junior/Entry"
salary_df$experience_level[salary_df$experience_level == "EX"] <- "Director/Executive"

#Renamed abbreviations for the employment_type column
salary_df$employment_type[salary_df$employment_type == "FT"] <- "FullTime"
salary_df$employment_type[salary_df$employment_type == "PT"] <- "PartTime"
salary_df$employment_type[salary_df$employment_type == "CT"] <- "Contract"
salary_df$employment_type[salary_df$employment_type == "FL"] <- "FreeLance"

#Renamed abbreviations for the company_size column
salary_df$company_size[salary_df$company_size == "S"] <- "Small"
salary_df$company_size[salary_df$company_size == "M"] <- "Medium"
salary_df$company_size[salary_df$company_size == "L"] <- "Large"

#Pulling columns from salary_df and iso into vectors that we can iterate from.
employee_residence <- salary_df %>% pull(employee_residence)

Alpha.2.code <- iso_df %>% pull(Alpha.2.code)

Country_name <- iso_df %>% pull(English.short.name.lower.case)

index <- 0
new_employee_residence <- c()

#Looping through each item in new_employee_residence
for (item in employee_residence) {
  
#check where the item exists in Alpha.2.code, then assign the index found in Alpha.2.code to index
    index <- which(Alpha.2.code == item)[1]
    
#use the index to find the corresponding Country_name, then append that Country_name to new_employee_residence, save to new_employee_residence
    new_employee_residence <- append(new_employee_residence, Country_name[index])
}

# assign new_employee_residence to the employee_residence column
salary_df$employee_residence <- new_employee_residence
```

```{r}
str(salary_df$experience_level)
str(salary_df$employment_type)
str(salary_df$employee_residence)
str(salary_df$company_size)
str(salary_df$company_location)
```

2. We do not need the salary or salary_currency columns for our analysis.¶
```{r}
salary_df <- salary_df %>%
  select(-c(X, salary, salary_currency))

str(salary_df)
```

3. Since remote_ratio contains ratios they all should have % at the end.¶

```{r}
salary_df$remote_ratio <- gsub("%", "", salary_df$remote_ratio)

```

```{r}
#Add a % at the end of every element in remote_ratio
remote_ratio <- salary_df %>% pull(remote_ratio)
new_remote_ratio <- paste0(remote_ratio, "%")

salary_df$remote_ratio <- new_remote_ratio

```

```{r}
str(salary_df$remote_ratio)
```

Analysis

```{r Analysis}
library(ggplot2)
library(dplyr)

experience_level <- data.frame(experience_level = salary_df$experience_level)
experience_counts <- experience_level %>% count(experience_level)

ggplot(data = experience_level) +
  geom_bar(mapping = aes(x = experience_level, fill = experience_level)) +
  geom_text(data = experience_counts, aes(x = experience_level, y = n, label = n), vjust = -0.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), text = element_text(size = 12)) +
  labs(title = "Distribution of Experience Level",
       x = "Experience Level",
       y = "Count",
       fill = "Experience Level") +
  coord_cartesian(ylim = c(0, 300))

```
Most people with data science related jobs are Senior level.

```{r}
library(ggplot2)
library(tidyr)

job_titles <- c("Data Engineer", "Data Scientist", "Data Analyst", "Machine Learning Engineer", "Analytics Engineer", "Data Architect", "Research Scientist", "Applied Scientist", "Data Science Manager", "Research Engineer")

query <- salary_df %>%
  filter(job_title %in% job_titles) %>%
  group_by(job_title, experience_level) %>%
  summarise(Count = n(), .groups = "drop") %>%
  complete(job_title = job_titles, experience_level, fill = list(Count = 0)) %>%
  arrange(desc(Count))

ggplot(query, aes(x = reorder(job_title, -Count), y = Count, fill = experience_level)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = Count, group = experience_level), position = position_dodge2(preserve = "single", width = 0.9), vjust = -0.5, size = 2.5, color = "black") +
  scale_x_discrete(labels = job_titles) +
  theme(axis.text.x = element_text(angle = 50, hjust = 1, vjust = 1), text = element_text(size = 10)) +
  labs(title = "Experience Level / Job Title", subtitle = "For the 10 most popular data science jobs", x = "Job Title", y = "Count") +
  guides(fill = guide_legend(title = "Experience Level")) +
  coord_flip() +
  coord_cartesian(ylim = c(0, 80))

```
The ratio of Seniors vs other experience levels is roughly the same per job title.

```{r}
library(ggplot2)

employment_type <- data.frame(employment_type = salary_df$employment_type)

ggplot(data = employment_type, aes(x = employment_type)) +
  geom_bar(fill = "orange") +
  geom_text(stat = "count", aes(label = after_stat(count)), vjust = -0.5, color = "black") +
  labs(title = "Employment Type Distribution", x = "Employment Type", y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), text = element_text(size = 14)) + 
  scale_y_continuous(limits = c(0, 700))

```
Virtually all employees are full time.

```{r}
job_title_counts <- table(salary_df$job_title)

# Order the job titles in decreasing order
ordered_counts <- sort(job_title_counts, decreasing = TRUE)

# Display the ordered job title counts
ordered_counts

```
```{r}
library(wordcloud)
## Loading required package: RColorBrewer
library(RColorBrewer)

# Convert job_title to character
salary_df$job_title <- as.character(salary_df$job_title)

# Filter the data based on job titles with count > 5
filtered_counts <- job_title_counts[job_title_counts > 5]

# Generate 20 bright colors by repeating the "Set1" palette
num_colors <- 20
colors <- rep(brewer.pal(9, "Set1"), length.out = num_colors)

# Create a data frame with word frequencies
word_data <- data.frame(word = names(filtered_counts),
                        freq = as.numeric(filtered_counts))

# Set a dark background for the word cloud
par(bg = "black")

# Generate the word cloud using wordcloud
set.seed(123)  # For reproducibility of colors and angles
wordcloud(word_data$word, word_data$freq, scale = c(2.5, 1), random.order = TRUE,
          colors = colors, rot.per = 0.2, random.color = TRUE)

```


```{r}
library(ggplot2)
library(dplyr)

# Create a data frame with salary counts
salary <- salary_df %>%
  group_by(salary_in_usd) %>%
  summarise(count = n())

# Convert salary_in_usd to numeric
salary$salary_in_usd <- as.numeric(as.character(salary$salary_in_usd))

# Sort the data frame by salary_in_usd
salary <- salary[order(salary$salary_in_usd), ]

# Set plot dimensions
options(repr.plot.width = 15, repr.plot.height = 15)

# Create the histogram plot with binwidth of 100000
ggplot(salary, aes(x = salary_in_usd)) +
  geom_histogram(binwidth = 100000, fill = "lightblue", color = "black") +
  scale_x_continuous(labels = scales::comma) +
  theme(text = element_text(size = 20)) +
  labs(title = "Distribution of Salary", x = "Salary", y = "Count")

# Print summary statistics
print(summary(salary))
```


```{r}
library(ggplot2)
library(scales)

salary_df$salary_in_usd <- as.numeric(salary_df$salary_in_usd)

temp <- data.frame(table(salary_df$job_title))

temp <- temp %>%
  arrange(desc(Freq)) %>%
  head(n = 70)

data <- salary_df %>%
  select(job_title, salary_in_usd) %>%
  filter(job_title == "Data Engineer" | job_title == "Data Scientist" | job_title == "Data Analyst" | job_title == "Machine Learning Engineer" | job_title == "Analytics Engineer" | job_title == "Data Architect" | job_title == "Research Scientist" | job_title == "Applied Scientist" | job_title == "Data Science Manager" | job_title == "Research Engineer" | job_title == "ML Engineer" | job_title == "Data Manager" | job_title == "Machine Learning Scientist" | job_title == "Data Science Consultant" | job_title == "Data Analytics Manager" | job_title == "Computer Vision Engineer" | job_title == "AI Scientist" | job_title == "BI Data Analyst" | job_title == "Business Data Analyst" | job_title == "Data Specialist" | job_title == "BI Developer" | job_title == "Applied Machine Learning Scientist" | job_title == "AI Developer" | job_title == "Big Data Engineer" | job_title == "Director of Data Science" | job_title == "Machine Learning Infrastructure Engineer" | job_title == "Applied Data Scientist" | job_title == "Data Operations Engineer" | job_title == "ETL Developer" | job_title == "Head of Data")

options(repr.plot.width = 15, repr.plot.height = 15)

data %>%
     ggplot(aes(job_title, salary_in_usd, fill = job_title))+
     geom_boxplot()+
     theme_bw()+
     theme(axis.text.x = element_text(angle = 90), legend.position = "none", text = element_text(size = 10))+
     scale_y_continuous(labels = comma)+
     labs(title = "Salary Range Based on Job Title for Data Science Jobs", subtitle = "30 most popular data science jobs", x = "Job Titles", y = "Salary (USD)")
```

```{r}
library(ggplot2)
library(magrittr)

remote_ratio <- data.frame(table(salary_df$remote_ratio))

options(repr.plot.width = 15, repr.plot.height = 15)

remote_ratio %>%
  ggplot() +
  aes(x = Var1, y = Freq) +
  geom_col(fill = "lightgreen", color = "black") +
  geom_text(aes(label = Freq), vjust = -0.5, size = 5) +
  xlab("Remote Ratio") +
  ylab("Count") +
  ylim(c(0, 500)) +  # Set y-axis limits (adjust the upper limit as desired)
  theme(text = element_text(size = 16)) +
  labs(title = "Distribution of Remote Ratio") +
  theme(plot.margin = margin(10, 10, 10, 10, "pt"))
```


```{r}
company_size <- data.frame(table(salary_df$company_size))

options(repr.plot.width = 15, repr.plot.height = 15)

 company_size %>%
    ggplot() +
    aes(x = Var1, y = Freq) +
    geom_col(fill = "purple", color = "black") +
    geom_text(aes(label = Freq), vjust = -0.5, size = 5) +
    labs(title = "Company Size Distribution", x = "Company Size", y = "Count") +
    ylim(c(0, 450)) +  # Set y-axis limits (adjust the upper limit as desired)
    theme(text = element_text(size = 16)) +
    theme(plot.margin = margin(10, 10, 10, 10, "pt"))
```
Conclusion:
In this analysis, I explored the salaries of data science jobs in the US. I identified some trends and patterns in the data science job market, such as most employee’s experience being senior-level and most data science companies are located in the US with company sizes ranging from 50-250 employees. In conclusion, my analysis provides a valuable snapshot of the current state and trends of data science salaries in the US, which can help data scientists and employers make informed decisions and plan their careers.
